
## Read in Data
* Map column values according to data dictionary
* Clean column names

```{r}
library(plyr)
library(tidyverse)
library(ggplot2)
library(gtools)

raw_df <- read.csv("../data/SY18-19_all_variables_Data.csv", check.names = FALSE)
df <- raw_df
value_mapping <- read.csv("../data/SY18-19_all_variables_ValueLabels.csv")

# convert the sector into the actual names
for (i in unique(value_mapping$VariableName)){
  temp_map_val <- value_mapping %>% subset(VariableName == i)
  df[[i]] <- mapvalues(df[[i]], from = temp_map_val$Value, to = temp_map_val$ValueLabel)
}

# clean up column names
colnames(df) <- gsub(" - ", "_", tolower(colnames(df)))
colnames(df) <- gsub("-|[[:space:]]+|/", "_", colnames(df))
colnames(df) <- gsub("\\(|\\)|\\:", "", colnames(df))
colnames(df) <- gsub("\\_(sfa1819|2018_19|drvgr2018|drvef2018|hd2018|2018_ef2018d|f1819|al2019|gasb_drvf2019|ic2018).*$", "", tolower(colnames(df)))
colnames(df) <- gsub("graduation_rate_bachelor_degree_within_6_years", "gradrate_ba_6yrs", colnames(df))

# combine the financial columns which were broken up for public and private schools into one
dupe_vals <- grep('value_of_endowment_assets_at_the_beginning_of_the_fiscal_year', colnames(df))
colnames(df)[dupe_vals[1]] <- 'value_of_endowment_assets_at_the_beginning_of_the_fiscal_year1'
colnames(df)[dupe_vals[2]] <- 'value_of_endowment_assets_at_the_beginning_of_the_fiscal_year2'

```


## Clean Data
* Combine financial columns
* Write Cleaned CSV file

### Combine financial columns
```{r}
df$endowment_total <- ifelse(is.na(df$value_of_endowment_assets_at_the_beginning_of_the_fiscal_year1),
                             df$value_of_endowment_assets_at_the_beginning_of_the_fiscal_year2,
                             df$value_of_endowment_assets_at_the_beginning_of_the_fiscal_year1)
df$finances_spent_research <- ifelse(is.na(df$research_current_year_total),
                                     df$research_total_amount,
                                     df$research_current_year_total)
df$finances_spent_student_services <- ifelse(is.na(df$student_service_total_amount),
                                             df$student_services_current_year_total,
                                             df$student_service_total_amount)
df$finances_spent_public_service <- ifelse(is.na(df$public_service_current_year_total),
                                           df$public_service_total_amount,
                                           df$public_service_current_year_total)
df$finances_spent_academic_support <- ifelse(is.na(df$academic_support_current_year_total),
                                             df$academic_support_total_amount,
                                             df$academic_support_current_year_total)
df$finances_spent_instruction <- ifelse(is.na(df$instruction_current_year_total),
                                        df$instruction_total_amount,
                                        df$instruction_current_year_total)
df$revenue_total <- ifelse(is.na(df$total_revenues_and_investment_return_total),
                           df$total_all_revenues_and_other_additions,
                           df$total_revenues_and_investment_return_total)

drop_cols <- c('value_of_endowment_assets_at_the_beginning_of_the_fiscal_year1', 'value_of_endowment_assets_at_the_beginning_of_the_fiscal_year2', 'research_current_year_total', 'research_total_amount', 'student_service_total_amount', 'student_services_current_year_total', 'public_service_current_year_total', 'public_service_total_amount', 'academic_support_current_year_total', 'academic_support_total_amount', 'instruction_current_year_total', 'instruction_total_amount', 'total_revenues_and_investment_return_total', 'total_all_revenues_and_other_additions', '')

df <- df[, !(colnames(df) %in% drop_cols)]
```

### Remove rows with missing values on enrollment and graduation rates
```{r}
# remove columns that have NA's
pre_df <- nrow(df)
print(paste("Total # of schools in dataset:", nrow(df)))

remove_any_nas <- c("undergraduate_enrollment",
                    "gradrate_ba_6yrs_total"
                    )

for (i in remove_any_nas){
  df <- df %>% subset(!(is.na(df[[i]]) | (df[[i]] == 0)))
}

# for each subgroup, if they have students from that group, they should also have a graduation rate
enroll_grad <- c("percent_of_undergraduate_enrollment_that_are_black_or_african_american" = "gradrate_ba_6yrs_black_non_hispanic",
                    "percent_of_undergraduate_enrollment_that_are_hispanic_latino" = "gradrate_ba_6yrs_hispanic",
                    "percent_of_undergraduate_enrollment_that_are_american_indian_or_alaska_native" = "gradrate_ba_6yrs_american_indian_or_alaska_native",
                    "percent_of_undergraduate_students_awarded_pell_grants" = "pell_grant_recipients_overall_graduation_rate_within_150_percent_of_normal_time"
                 )

for (i in names(enroll_grad)){
  df <- df %>% subset(ifelse(!is.na(df[[i]]) & (df[[i]] != 0),
                             !is.na(df[[enroll_grad[[i]]]]),
                             TRUE))
}


# remove any rows if they do not have *any* students from underrepresented groups
df <- df %>% mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>% mutate(sum_subgroups = percent_of_undergraduate_enrollment_that_are_black_or_african_american + percent_of_undergraduate_enrollment_that_are_hispanic_latino + percent_of_undergraduate_students_awarded_pell_grants + percent_of_undergraduate_enrollment_that_are_american_indian_or_alaska_native) %>% mutate(sum_race_subgroups = percent_of_undergraduate_enrollment_that_are_black_or_african_american + percent_of_undergraduate_enrollment_that_are_hispanic_latino + percent_of_undergraduate_enrollment_that_are_american_indian_or_alaska_native + percent_of_undergraduate_enrollment_that_are_native_hawaiian_or_other_pacific_islander)

# remove any schools where there are no underrepresented subgroups
df <- df %>% subset(sum_subgroups != 0)

# compute mean
df <- df %>% mutate(mean_subgroups = (percent_of_undergraduate_enrollment_that_are_black_or_african_american + percent_of_undergraduate_enrollment_that_are_hispanic_latino + percent_of_undergraduate_students_awarded_pell_grants + percent_of_undergraduate_enrollment_that_are_american_indian_or_alaska_native)/4)

# convert mean subgroups to quantiles
df <- df %>% mutate(diversity_quantiles = quantcut(df$mean_subgroups, q = 4, na.rm = TRUE))
df$diversity_quantiles <- mapvalues(df$diversity_quantiles,
                                    from = sort(unique(quantcut(df$mean_subgroups, q = 4, na.rm = TRUE))),
                                    to = paste("diversity quantile", 1:4))

# convert the sector into the actual names
for (i in unique(value_mapping$VariableName)){
  temp_map_val <- value_mapping %>% subset(VariableName == i)
  df[[i]] <- mapvalues(df[[i]], from = temp_map_val$Value, to = temp_map_val$ValueLabel)
}


paste("Removed", pre_df-nrow(df), "rows with missing undergraduate enrollment and graduation rates data.")
paste("Total # of schools for analysis:", nrow(df))

write.csv(df, "../data/all_data_merged_df(NEW).csv", row.names = FALSE)

```

## Perform EDA

#### Grab Georgetown Values
```{r}
gtown <- df %>% subset(institution_name == 'Georgetown University')
gtown
```


### Underrepresented Students Distribution
```{r}
# mean of underrepresneted group makeup of the populations size histogram
# df %>% ggplot(aes(x=mean_subgroups)) + geom_boxplot(fill = "#bca0dc") + ggtitle("Percent of Undergraduate Population Belonging to\nan Underrepresented Demographic*\nfor all 4-Year Colleges in the United States") + xlab("% of Undergraduate Population in an Underrepresented Demographics") + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$mean_subgroups, color='#000080', linetype = "dashed")

# sum of all underrepresented races = black, hispanic, native american, and pacific islander 
df %>% ggplot(aes(x=sum_race_subgroups)) + geom_boxplot(fill = "#bca0dc") + ggtitle("Percent of Undergraduate Population Belonging to\nan Underrepresented Race/Ethnicity Demographic*\nfor all 4-Year Colleges in the United States") + xlab("% of Undergraduate Population in an Underrepresented Race/Ethnicity*") + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$sum_race_subgroups, color='#000080', linetype = "dashed")


# does racial diversity lead to overall higher graduation rates?
ggplot(df, aes(y=gradrate_ba_6yrs_total, x=sum_race_subgroups, color = sector_of_institution)) + 
  geom_point(aes(size=undergraduate_enrollment, fill = sector_of_institution), colour="white",pch=21) + geom_vline(xintercept=median(df$sum_race_subgroups), linetype='dashed', color='red') + geom_hline(yintercept=median(df$gradrate_ba_6yrs_total), linetype='dashed', color='red') + labs(y='College\'s 6-Year Graduation Rate (%)', x='% of Undergraduate Population in\nUnderrepresented Racial Category') + ggtitle("Overall Graduation Rates and\nDiverse Makeup of Undergraduate Population") + theme_gray() +  theme(plot.title = element_text(hjust = 0.5)) + geom_point(aes(gtown$gradrate_ba_6yrs_total, gtown$sum_race_subgroups), fill='#000080', colour="red", pch=21)
# geom_boxplot(aes(x=variable, y=value, fill=variable))

head(df %>% arrange(desc(sum_subgroups)) %>% select(institution_name, sum_subgroups))
```

### Size of College (by Undergraduate Enrollment)
```{r}
# enrollemnt size histogram
df %>% ggplot(aes(x=undergraduate_enrollment)) + geom_histogram(position="stack", col = "white", fill = "#bca0dc") + ggtitle("Undergraduate Enrollment Numbers for\nAll 4-Year Colleges in the United States") + xlab("Number of Undergraduates Enrolled") + ylab("# of Colleges") + geom_vline(xintercept=median(df$undergraduate_enrollment), color='#b30000') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$undergraduate_enrollment, color='#000080', linetype = "dashed")
```


#### Look into Outliers
* After googling each of the largest schools, we see that their abnormal size is due to being primarily online institutions. We decided to keep these institutions because it would be informative to know if primarily online institutions graduate underrepresented students at higher rates. Decided to take the logarithm of the enrollment size in order to coerce the distribution into being more normal

```{r}
# look into outliers
df %>% subset(undergraduate_enrollment > 40000) %>% select(institution_name, undergraduate_enrollment, all_programs_offered_completely_via_distance_education, percent_of_undergraduate_students_enrolled_exclusively_in_distance_education_courses) %>% arrange(desc(undergraduate_enrollment))
# found out that these are all online colleges. question - would we want to include entirely online colleges? we should at least pull in that data point
```


```{r}
library(scales)
# redo histogram by taking log enrollment
df %>% ggplot(aes(x=undergraduate_enrollment)) + scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) + geom_histogram(position="stack", col = "white", fill = "#bca0dc") + ggtitle("Undergraduate Enrollment Numbers for\nAll 4-Year Colleges in the United States") + xlab("Number of Undergraduates Enrolled") + ylab("# of Colleges") + geom_vline(xintercept=mean(df$undergraduate_enrollment), color='#b30000') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$undergraduate_enrollment, color='#000080', linetype = "dashed")

```


### 6-Year Graduation Rates
```{r}
# graduation rates histogram
df %>% ggplot(aes(x=gradrate_ba_6yrs_total)) + geom_histogram(position="stack", col = "white", fill = "#a3c585", bins = 10) + ggtitle("6-Year Graduation Rates for\nAll 4-Year Colleges in the United States") + xlab("6-Year Graduation Rates (%)") + ylab("# of Colleges") + geom_vline(xintercept=mean(df$gradrate_ba_6yrs_total), color='#b30000') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$gradrate_ba_6yrs_total, color='#000080', linetype = "dashed")

```
#### Look into Outliers
```{r}
# look out graduate rate outliers

df %>% subset(gradrate_ba_6yrs_total < 10) %>% select(institution_name, undergraduate_enrollment, sector_of_institution, institutional_category, gradrate_ba_6yrs_total, gradrate_ba_6yrs_black_non_hispanic, gradrate_ba_6yrs_white_non_hispanic)
```


```{r}
library(reshape2)
library(ggplot2)

ggplot(melt(rename(df, Men = gradrate_ba_6yrs_men, Women = gradrate_ba_6yrs_women), id.vars = 'unitid', measure.vars = c('Men', 'Women'))) + geom_boxplot(aes(x=variable, y=value, fill=variable)) + ggtitle("Comparing 6-Year Graduation Rates for Each Sex")  + xlab("Gender") + ylab("6-Year Graduation Rates (%)") +  geom_hline(yintercept=mean(df$gradrate_ba_6yrs_total), color='#b30000', linetype ='dashed') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + scale_fill_discrete(name = 'Gender') + geom_hline(yintercept=gtown$gradrate_ba_6yrs_total, color='#000080', linetype = "dashed")

```

## Comparing Graduation Rates for Different Subgroups of Students
```{r}
# any points above the red dotted lines are schools where Pell-Grant recipients are graduating at *higher* rates than the general student body
# we added the dimension of school size to the mix and as you can see, majority of the schools where Pell-Grant recipients are graduating at much higher rates than the general student body are smaller sized schools
ggplot(df, aes(x=gradrate_ba_6yrs_total, y=pell_grant_recipients_overall_graduation_rate_within_150_percent_of_normal_time, color = sector_of_institution)) + 
  geom_point(aes(size=undergraduate_enrollment, fill = sector_of_institution), colour="white",pch=21) + geom_abline(intercept=0, slope=1, linetype='dashed', color='red') + labs(y='Pell Recipients 6-Year Graduation Rate (%)', x='6-Year Graduation Rate All Students (%)') + ggtitle("Comparing Overall Graduation Rates to\nPell Recipients Graduation Rates") + theme_gray() +  theme(plot.title = element_text(hjust = 0.5)) + geom_point(aes(gtown$gradrate_ba_6yrs_total, gtown$pell_grant_recipients_overall_graduation_rate_within_150_percent_of_normal_time), fill='#000080', colour="red", pch=21)

paste("There are", sum(df$pell_grant_recipients_overall_graduation_rate_within_150_percent_of_normal_time > df$gradrate_ba_6yrs_total), "schools where Pell-Grant recipients graduate at higher rates than that school's overall population.")

# create a derived field that indicates whether pell students are doing about the same or better than overall student body
df <- df %>% mutate(pell_above_avg = pell_grant_recipients_overall_graduation_rate_within_150_percent_of_normal_time >= gradrate_ba_6yrs_total)
```

```{r}
# any points above the red dotted lines are schools where women graduate at higher rates than men
ggplot(df, aes(y=gradrate_ba_6yrs_women, x=gradrate_ba_6yrs_men, color = sector_of_institution)) + 
  geom_point(aes(size=undergraduate_enrollment, fill = sector_of_institution), colour="white",pch=21) + geom_abline(intercept=0, slope=1, linetype='dashed', color='red') + labs(x='6-Year Graduation Rate for Men', y='6-Year Graduation Rate for Women') + ggtitle("Comparing Graduation Rates of Men vs. Women") + theme_gray() +  theme(plot.title = element_text(hjust = 0.5)) + geom_point(aes(gtown$gradrate_ba_6yrs_women, gtown$gradrate_ba_6yrs_men), fill='#000080', colour="red", pch=21)

# create a derived field that indicates whether women and men are graduating at the same rates
df <- df %>% mutate(women_above_avg = gradrate_ba_6yrs_women >= gradrate_ba_6yrs_men)
```

```{r}
# any points above the red dotted lines are schools where black students graduate at higher rates than white students
ggplot(df, aes(y=gradrate_ba_6yrs_black_non_hispanic, x=gradrate_ba_6yrs_total)) + 
  geom_point(aes(size=undergraduate_enrollment, fill = sector_of_institution), colour="white", pch=21) + geom_abline(intercept=0, slope=1, linetype='dashed', color='red') + labs(x='6-Year Graduation Rate for All Students', y='6-Year Graduation Rate for Black Non-Hispanic Students') + ggtitle("Comparing Black Students' Graduation Rates\nto Each College's Total Graduation Rate") + theme_gray() +  theme(plot.title = element_text(hjust = 0.5)) + geom_point(aes(gtown$gradrate_ba_6yrs_black_non_hispanic, gtown$gradrate_ba_6yrs_total), fill='#000080', colour="red", pch=21)

paste("There are", sum(df$gradrate_ba_6yrs_black_non_hispanic > df$gradrate_ba_6yrs_total), "schools where Black students graduate at higher rates than White students.")

# create a derived field that indicates whether black students are graduating at the same rates at the school avg
df <- df %>% mutate(black_above_avg = gradrate_ba_6yrs_black_non_hispanic >= gradrate_ba_6yrs_total)
```

```{r}
# any points above the red dotted lines are schools where hispanic students graduate at higher rates than white students
ggplot(df, aes(y=gradrate_ba_6yrs_hispanic, x=gradrate_ba_6yrs_total, color = sector_of_institution)) + 
  geom_point(aes(size=undergraduate_enrollment, fill = sector_of_institution), colour="white",pch=21) + geom_abline(intercept=0, slope=1, linetype='dashed', color='red') + labs(x='6-Year Graduation Rate for All Students', y='6-Year Graduation Rate for Hispanic Students') + ggtitle("Comparing Hispanic Students' Graduation Rates\nto Each College's Total Graduation Rate") + theme_gray() +  theme(plot.title = element_text(hjust = 0.5)) + geom_point(aes(gtown$gradrate_ba_6yrs_hispanic, gtown$gradrate_ba_6yrs_total), fill='#000080', colour="red", pch=21)

paste("There are", sum(df$gradrate_ba_6yrs_hispanic > df$gradrate_ba_6yrs_total), "schools where Hispanic/Latinx students graduate at higher rates than White students.")


# create a derived field that indicates whether hispanic students are graduating at the same rates at the school avg
df <- df %>% mutate(hispanic_above_avg = gradrate_ba_6yrs_hispanic >= gradrate_ba_6yrs_total)
```


```{r}

hist(df$gradrate_ba_6yrs_american_indian_or_alaska_native)
print(sum(!is.na(df$gradrate_ba_6yrs_american_indian_or_alaska_native)))

# any points above the red dotted lines are schools where hispanic students graduate at higher rates than white students
ggplot(df, aes(y=gradrate_ba_6yrs_american_indian_or_alaska_native, x=gradrate_ba_6yrs_total, color = sector_of_institution)) + geom_point(aes(size=undergraduate_enrollment, fill = sector_of_institution), colour="white",pch=21) + geom_abline(intercept=0, slope=1, linetype='dashed', color='red') + labs(x='6-Year Graduation Rate for White Non-Hispanic Students', y='6-Year Graduation Rate for\nAmerican Indian or Alaska Native Students (%)') + ggtitle("Comparing Native American Students'\nGraduation Rates to Each College's\nTotal Graduation Rate (%)") + theme_gray() +  theme(plot.title = element_text(hjust = 0.5)) + geom_point(aes(gtown$gradrate_ba_6yrs_american_indian_or_alaska_native, gtown$gradrate_ba_6yrs_total), fill='#000080', colour="red", pch=21)


paste("There are", sum(df$gradrate_ba_6yrs_american_indian_or_alaska_native > df$gradrate_ba_6yrs_total), "schools where Native American students graduate at higher rates than White students.")


```
  
## Dive into Schools where Underrepresented Student Success Rates Mirror or Surpass General Student Body
```{r}
# see how many schools are ones in which underrepresented students are doing the same or better as the general student body
df <- df %>% mutate(underrepresented_above_avg = hispanic_above_avg & pell_above_avg & women_above_avg & black_above_avg)

print(table(df$underrepresented_above_avg))

subgroups_success_df <- df %>% subset(underrepresented_above_avg==TRUE)

# df %>% subset(underrepresented_above_avg == TRUE) %>% ggplot(aes(x=gradrate_ba_6yrs_total,fill = underrepresented_above_avg)) + geom_histogram(position="stack", col = "white") + ggtitle("Seeing Graduation Rates for Schools\nWhere Underrepresented Demographics Perform\nin par w/ Rest of Student Body") + xlab("Graduation Rate") + ylab("# of Colleges") + geom_vline(xintercept=median(df$gradrate_ba_6yrs_total), color='#b30000') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$gradrate_ba_6yrs_total, color='#000080', linetype = "dashed")


df %>% ggplot(aes(x=underrepresented_above_avg, fill = underrepresented_above_avg)) + geom_bar(position="stack", col = "white") + ggtitle("Seeing Graduation Rates for Schools\nWhere Underrepresented Demographics Perform\nin par w/ Rest of Student Body") + xlab("Graduation Rate") + ylab("# of Colleges") + geom_vline(xintercept=median(df$gradrate_ba_6yrs_total), color='#b30000') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$gradrate_ba_6yrs_total, color='#000080', linetype = "dashed")

df %>% ggplot(aes(x=hispanic_above_avg, fill = hispanic_above_avg)) + geom_bar(position="stack", col = "white") + ggtitle("Seeing Graduation Rates for Schools\nWhere Hispanic/Latinx Students Perform\nin par w/ Rest of Student Body") + xlab("Graduation Rate") + ylab("# of Colleges") + geom_vline(xintercept=median(df$gradrate_ba_6yrs_total), color='#b30000') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$gradrate_ba_6yrs_total, color='#000080', linetype = "dashed")

df %>% ggplot(aes(x=black_above_avg, fill = black_above_avg)) + geom_bar(position="stack", col = "white") + ggtitle("Seeing Graduation Rates for Schools\nWhere Black Students Perform\nin par w/ Rest of Student Body") + xlab("Graduation Rate") + ylab("# of Colleges") + geom_vline(xintercept=median(df$gradrate_ba_6yrs_total), color='#b30000') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$gradrate_ba_6yrs_total, color='#000080', linetype = "dashed")

df %>% ggplot(aes(x=pell_above_avg, fill = pell_above_avg)) + geom_bar(position="stack", col = "white") + ggtitle("Seeing Graduation Rates for Schools\nWhere Low-Income Students Perform\nin par w/ Rest of Student Body") + xlab("Graduation Rate") + ylab("# of Colleges") + geom_vline(xintercept=median(df$gradrate_ba_6yrs_total), color='#b30000') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$gradrate_ba_6yrs_total, color='#000080', linetype = "dashed")
```

```{r}
subgroups_success_df %>% ggplot(aes(x=sector_of_institution, fill = sector_of_institution)) + geom_bar(position="stack", col = "white") + ggtitle("Sector of Institution") + xlab("Graduation Rate") + ylab("# of Colleges") + geom_vline(xintercept=median(df$gradrate_ba_6yrs_total), color='#b30000') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$gradrate_ba_6yrs_total, color='#000080', linetype = "dashed")

subgroups_success_df %>% ggplot(aes(x=undergraduate_enrollment)) + scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) + geom_histogram(position="stack", col = "white", fill = "#bca0dc", bins=5) + ggtitle("Undergraduate Enrollment Numbers for\nAll 4-Year Colleges in the United States") + xlab("Number of Undergraduates Enrolled") + ylab("# of Colleges") + geom_vline(xintercept=mean(df$undergraduate_enrollment), color='#b30000') + theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(xintercept=gtown$undergraduate_enrollment, color='#000080', linetype = "dashed")
```
## Do schools where underrepresented students do better have higher graduation rates?
```{r}
df %>%
  ggplot( aes(x = underrepresented_above_avg, y = gradrate_ba_6yrs_total, fill=underrepresented_above_avg)) +
  geom_boxplot() +ggtitle("Comparing Graduation Rates at Institutions where\nUnderrepresented Students Succeed")


df %>%
  ggplot( aes(x = hispanic_above_avg, y = gradrate_ba_6yrs_total, fill=hispanic_above_avg)) +
  geom_boxplot() +ggtitle("Comparing Graduation Rates at Institutions where\nHispanic/Latinx Students Succeed")

df %>%
  ggplot( aes(x = black_above_avg, y = gradrate_ba_6yrs_total, fill=black_above_avg)) +
  geom_boxplot() +ggtitle("Comparing Graduation Rates at Institutions where\nBlack Students Succeed")

df %>%
  ggplot( aes(x = pell_above_avg, y = gradrate_ba_6yrs_total, fill=pell_above_avg)) +
  geom_boxplot() +ggtitle("Comparing Graduation Rates at Institutions where\nLow-Income Students Succeed")
```
## Is a factor of whether Low-Income students succeed percentage of costs covered by grants?
```{r}
# avg amount of federal, state, local, institutional sums up all institutional aid a student can get
# total price for in state tution, off campus. in state vs. out of state typically only counts for public schools. low-income students are highly unlikely to attend an out of state public school
df <- df %>% mutate(cost_metric = ifelse(is.na(total_price_for_in_state_students_living_off_campus_not_with_family) | total_price_for_in_state_students_living_off_campus_not_with_family == 0,
                                         total_price_for_in_state_students_living_on_campus,
                                         total_price_for_in_state_students_living_off_campus_not_with_family))

df <- df %>% mutate(percent_covered_aid = average_amount_of_federal_state_local_institutional_or_other_sources_of_grant_aid_awarded_to_undergraduate_students/cost_metric) %>% mutate(percent_covered_aid = ifelse(cost_metric==0, 1, percent_covered_aid))

hist(df$percent_covered_aid)
boxplot(df$percent_covered_aid)

pell_df <- df %>% subset(pell_above_avg==TRUE)


nrow(pell_df)

hist(df$average_amount_pell_grant_aid_awarded_to_undergraduate_students)
hist(df$percent_of_undergraduate_students_awarded_pell_grants)
boxplot(df$average_net_price_income_0_30_000_students_awarded_title_iv_federal_financial_aid)
```

## Hypothesis test: Financial Aid Coverage <> Low-Income Students' Grad Rates
```{r}
mean_aid <- mean(df$percent_covered_aid, na.rm=TRUE)
df <- df %>% mutate(above_avg_aid = ifelse(percent_covered_aid > mean_aid, "above avg. aid", "avg. aid or below"))
print(table(df$above_avg_aid))

# get georgetown value
gtown <- df %>% subset(institution_name == 'Georgetown University')
print(gtown$percent_covered_aid)

df %>%
  ggplot( aes(x = above_avg_aid, y = pell_grant_recipients_overall_graduation_rate_within_150_percent_of_normal_time, fill=above_avg_aid)) +
  geom_boxplot() + ggtitle("Low-Income Grad Rates at Schools w/\nDifferent Financial Aid Coverage")

# null hypothesis: the average graduation rate for pell students at schools where the average aid given covers more than the median amount is the same as schools where average aid is median amount or less
# alternative hypothesis: the average graduation rate for pell students at schools where the average aid given covers more than the median amount is MORE than schools where average aid is median amount or less
t.test(pell_grant_recipients_overall_graduation_rate_within_150_percent_of_normal_time ~ above_avg_aid,
       data = df,
       alternative = "greater")
```

## Hypothesis test: Diversity <> Hispanic/Latinx' Grad Rates
```{r}
(mean_latinx <- mean(df$percent_of_undergraduate_enrollment_that_are_hispanic_latino, na.rm=TRUE))

df <- df %>% mutate(diverse_latinx =  ifelse(percent_of_undergraduate_enrollment_that_are_hispanic_latino > mean_latinx,
                                               'latinx diverse',
                                               'not latinx diverse'))

df %>%
  ggplot( aes(x = diverse_latinx, y = gradrate_ba_6yrs_hispanic, fill=diverse_latinx)) +
  geom_boxplot() + ggtitle("Hispanic/Latinx Grad Rates at Schools w/\nDifferent Hispanic/Latinx Populations")


# null hypothesis: the average graduation rate for latinx students at schools where the demographic is above avg. latinx population is the same as schools where latinx population is avg
# alternative hypothesis: the average graduation rate for latinx students at schools where the demographic is above avg. latinx population is MORE THAN at schools where latinx population is avg
t.test(gradrate_ba_6yrs_hispanic ~ diverse_latinx,
       data = df,
       alternative = "greater")
```



## Hypothesis test: Diversity <> Black Grad Rates
```{r}
(mean_black <- mean(df$percent_of_undergraduate_enrollment_that_are_black_or_african_american, na.rm=TRUE))

df <- df %>% mutate(diverse_black =  ifelse(percent_of_undergraduate_enrollment_that_are_black_or_african_american > mean_black,
                                               'black/aa diverse',
                                               'not black/aa diverse'))
print(table(df$diverse_black))

df %>%
  ggplot( aes(x = diverse_black, y = gradrate_ba_6yrs_black_non_hispanic, fill=diverse_black)) +
  geom_boxplot() + ggtitle("Black/African American Grad Rates at Schools w/\nDifferent Black/African American Populations")


# null hypothesis: the average graduation rate for black / african american students at schools where the demographic is above avg. black / african american population is the same as schools where black / african american population is avg
# alternative hypothesis: the average graduation rate for black / african american students at schools where the demographic is above avg. black / african american population is MORE THAN at schools where black / african american population is avg
t.test(gradrate_ba_6yrs_black_non_hispanic ~ diverse_black,
       data = df,
       alternative = "less")
```



## Hypothesis test: Diversity <> Native American Grad Rates
```{r}
(mean_native <- mean(df$percent_of_undergraduate_enrollment_that_are_american_indian_or_alaska_native, na.rm=TRUE))

df <- df %>% mutate(diverse_native =  ifelse(percent_of_undergraduate_enrollment_that_are_american_indian_or_alaska_native > mean_native,
                                               'american indian/native diverse',
                                               'not american indian/native diverse'))

print(table(df$diverse_native))

df %>%
  ggplot( aes(x = diverse_native, y = gradrate_ba_6yrs_american_indian_or_alaska_native, fill=diverse_native)) +
  geom_boxplot() + ggtitle("American Indian/Native American Grad Rates at Schools w/\nDifferent American Indian/Native American Populations")

# null hypothesis: the average graduation rate for american indian/native american students at schools where the demographic is above avg. american indian/native american population is the same as schools where american indian/native american population is avg
# alternative hypothesis: the average graduation rate for american indian/native american students at schools where the demographic is above avg. american indian/native american population is MORE THAN at schools where american indian/native american population is avg
t.test(gradrate_ba_6yrs_american_indian_or_alaska_native ~ diverse_native,
       data = df,
       alternative = "greater")
```