---
title: "Single and Multiple Linear Regression"
author: "Valeria Vera"
date: "2022-12-02"
output: html_document
---

```{r}
df <- read.csv('../data/SY18-19_all_variables_Data.csv')
library(tidyverse)
library(ggpubr)
theme_set(theme_pubr())
library(ISLR2)
library(leaps)

library(tidyverse)
library(caret)
```


Multiple linear regression

```{r}


#renaming columns
colnames(df)[which(names(df) == "Graduation.rate...Bachelor.degree.within.6.years..women..DRVGR2018.")] <- "gradrate_women"
colnames(df)[which(names(df) == "Graduation.rate...Bachelor.degree.within.6.years..men..DRVGR2018.")] <- "gradrate_men"
colnames(df)[which(names(df) == "Graduation.rate...Bachelor.degree.within.6.years..total..DRVGR2018.")] <- "gradrate_all"

colnames(df)[which(names(df) == "Percent.of.undergraduate.students.awarded.federal..state..local..institutional.or.other.sources.of.grant.aid..SFA1819.")] <- "students_percent_aid"
colnames(df)[which(names(df) == "Percent.of.undergraduate.students.awarded.Pell.grants..SFA1819.")] <- "students_percent_pellg"
colnames(df)[which(names(df) == "Average.amount.of.federal..state..local..institutional.or.other.sources.of.grant.aid.awarded.to.undergraduate.students..SFA1819.")] <- "students_avg_aid"
colnames(df)[which(names(df) == "Average.amount.Pell.grant.aid.awarded.to.undergraduate.students..SFA1819.")] <- "students_avg_pellg"
#colnames(df)[which(names(df) == "off_campus_not_with_family_room_and_board")] <- "living_exp_off_c"
#colnames(df)[which(names(df) == "on_campus_room_and_board")] <- "living_exp_on_c"

colnames(df)[which(names(df) == "Degree.of.urbanization..Urban.centric.locale...HD2018.")] <- "degree_urbanization"

colnames(df)[which(names(df) == "Research.Total.amount..F1819_F2.")] <- "amount_research"

df <- subset(df, select = c(gradrate_women,gradrate_men,gradrate_all,
                    students_percent_aid,students_percent_pellg,students_avg_aid,students_avg_pellg,degree_urbanization,amount_research))
df = df %>% drop_na()
df.scaled <- as.data.frame(scale(df))
```


```{r}
#outlier in amount_research


ggplot(df.scaled, aes(y=amount_research,x=  gradrate_women)) + 
geom_boxplot(outlier.colour="red",outlier.size=4)

quartiles <- quantile(df.scaled$amount_research, probs=c(.01, .995), na.rm = FALSE)
IQR <- IQR(df.scaled$amount_research)
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
data_no_outlier <- subset(df.scaled, df.scaled$amount_research < Upper)

ggplot(data_no_outlier, aes(y=amount_research,x=  gradrate_women)) + 
geom_boxplot(outlier.colour="red",outlier.size=4)

ggplot(df.scaled, aes(x=amount_research)) + 
    geom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.5,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666") +coord_cartesian(xlim=c(0,1),ylim=c(0,10))  # Overlay with transpar
#plotting amount research

y <- ggplot(data_no_outlier, aes(x = amount_research, y =  gradrate_women)) +
  geom_point(aes(color = amount_research),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and amount_research on campus") +
  xlab("amount_research") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = amount_research, y =  gradrate_men)) +
  geom_point(aes(color = amount_research),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and amount_research on campus") +
  xlab("amount_research") + ylab("Graduate rate of men")+ theme_gray() 
```

```{r}
y <- ggplot(df, aes(x = degree_urbanization, y =  gradrate_women)) +
  geom_point(aes(color = degree_urbanization),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and degree_urbanization") +
  xlab("degree_urbanization") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = degree_urbanization, y =  gradrate_men)) +
  geom_point(aes(color = degree_urbanization),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and degree_urbanization") +
  xlab("degree_urbanization") + ylab("Graduate rate of men")+ theme_gray() 

```

###Data Gone

```
y <- ggplot(df, aes(x = living_exp_on_c, y =  gradrate_women)) +
  geom_point(aes(color = living_exp_on_c),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and living expenses on campus") +
  xlab("living expenses by city (dollars)") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = living_exp_on_c, y =  gradrate_men)) +
  geom_point(aes(color = living_exp_on_c),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and living expenses on campus") +
  xlab("living expenses by city (dollars)") + ylab("Graduate rate of men")+ theme_gray() 

```

###Data Gone
```
y <- ggplot(df, aes(x = living_exp_on_c, y =  gradrate_women)) +
  geom_point(aes(color = living_exp_on_c),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and living expenses on campus") +
  xlab("living expenses by city (dollars)") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = living_exp_on_c, y =  gradrate_men)) +
  geom_point(aes(color = living_exp_on_c),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and living expenses on campus") +
  xlab("living expenses by city (dollars)") + ylab("Graduate rate of men")+ theme_gray() 

```


```
y <- ggplot(df, aes(x = living_exp_off_c, y =  gradrate_women)) +
  geom_point(aes(color = living_exp_off_c),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and living expenses out campus") +
  xlab("living expenses by city (dollars)") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = living_exp_off_c, y =  gradrate_men)) +
  geom_point(aes(color = living_exp_off_c),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and living expenses out campus") +
  xlab("living expenses by city (dollars)") + ylab("Graduate rate of men")+ theme_gray() 

```
```{r}
y <- ggplot(df, aes(x = students_percent_aid, y =  gradrate_women)) +
  geom_point(aes(color = students_percent_aid),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_percent_aid") +
  xlab("students_percent_aid") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = students_percent_aid, y =  gradrate_men)) +
  geom_point(aes(color = students_percent_aid),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_percent_aid") +
  xlab("students_percent_aid") + ylab("Graduate rate of men")+ theme_gray() 

```

```{r}
y <- ggplot(df, aes(x = students_percent_pellg, y =  gradrate_women)) +
  geom_point(aes(color = students_percent_pellg),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_percent_pellg") +
  xlab("students_percent_pellg") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = students_percent_pellg, y =  gradrate_men)) +
  geom_point(aes(color = students_percent_pellg),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_percent_pellg") +
  xlab("students_percent_pellg") + ylab("Graduate rate of men")+ theme_gray() 

```

```{r}
y <- ggplot(df, aes(x = students_avg_aid, y =  gradrate_women)) +
  geom_point(aes(color = students_avg_aid),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_avg_aid") +
  xlab("students_percent_pellg") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = students_avg_aid, y =  gradrate_men)) +
  geom_point(aes(color = students_avg_aid),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_avg_aid") +
  xlab("students_avg_aid") + ylab("Graduate rate of men")+ theme_gray() 

```

```{r}
y <- ggplot(df, aes(x = students_avg_pellg, y =  gradrate_women)) +
  geom_point(aes(color = students_avg_pellg),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_avg_pellg") +
  xlab("students_percent_pellg") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = students_avg_pellg, y =  gradrate_men)) +
  geom_point(aes(color = students_avg_pellg),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_avg_pellg") +
  xlab("students_avg_pellg") + ylab("Graduate rate of men")+ theme_gray() 

```


```{r}
#separating data
set.seed(1)
sample <- sample(c(TRUE, FALSE), nrow(df_clean), replace=TRUE, prob=c(0.8,0.2))
train  <- df[sample, ]
test   <- df[!sample, ]


#define and train model
model_m<- lm(gradrate_men ~ (students_percent_aid+students_percent_pellg+
                             students_avg_aid+students_avg_pellg+
                             +degree_urbanization+amount_research), data = train)
summary(model_m)

#evaluate model
coeff_table = summary(model_m)$coefficients
coeff_w = c(coeff_table[2,4], coeff_table[3,4],
coeff_table[4,4],coeff_table[5,4],coeff_table[6,4],coeff_table[7,4])
df_coeff_m = data.frame(coeff)
rownames(df_coeff_m) <- rownames(coeff_table)[-1]

predictions <- model_m %>% predict(test)
model_m_r= c(
error_rate = sigma(model_m)/mean(train$gradrate_men),
RMSE(predictions, test$gradrate_men),
R2(predictions, test$gradrate_men)
)
par(mfrow=c(2,2))
plot(model_m)
confint(model_m)
car::vif(model_m)
```
For a given the predictor, the t-statistic evaluates whether or not there is significant association between the predictor and the outcome variable, that is whether the beta coefficient of the predictor is significantly different from zero.

For a given predictor variable, the coefficient (b) can be interpreted as the average effect on y of a one unit increase in predictor, holding all other predictors fixed.


```{r}

#separating data
set.seed(1)
sample <- sample(c(TRUE, FALSE), nrow(df_clean), replace=TRUE, prob=c(0.8,0.2))
train  <- df[sample, ]
test   <- df[!sample, ]

#define and train model
model_w<- lm(gradrate_women ~ (students_percent_aid+students_percent_pellg+
                             students_avg_aid+students_avg_pellg+
                             +degree_urbanization+amount_research), data = train)
summary(model_w)
coeff_table = summary(model_w)$coefficients
coeff_w = c(coeff_table[2,4], coeff_table[3,4],
coeff_table[4,4],coeff_table[5,4],coeff_table[6,4],coeff_table[7,4])
df_coeff_w = data.frame(coeff)
rownames(df_coeff_w) <- rownames(coeff_table)[-1]

#evaluate model
predictions <- model_w %>% predict(test)
model_w_r= c(
error_rate = sigma(model_m)/mean(train$gradrate_women),
RMSE(predictions, test$gradrate_women),
R2(predictions, test$gradrate_women)
)
par(mfrow=c(2,2))
plot(model_w)
```

```{r}

#separating data
set.seed(1)
sample <- sample(c(TRUE, FALSE), nrow(df_clean), replace=TRUE, prob=c(0.8,0.2))
train  <- df[sample, ]
test   <- df[!sample, ]

#define and train model
model_a<- lm(gradrate_all ~ (students_percent_aid+students_percent_pellg+
                             students_avg_aid+students_avg_pellg+
                             +degree_urbanization+amount_research), data = train)
coeff_table = summary(model_a)$coefficients
coeff_w = c(coeff_table[2,4], coeff_table[3,4],
coeff_table[4,4],coeff_table[5,4],coeff_table[6,4],coeff_table[7,4])
df_coeff_a = data.frame(coeff)
rownames(df_coeff_a) <- rownames(coeff_table)[-1]

#evaluate model
predictions <- model_a %>% predict(test)
model_t_r= c(
error_rate = sigma(model_a)/mean(train$gradrate_all),
RMSE(predictions, test$gradrate_all),
R2(predictions, test$gradrate_all)
)
par(mfrow=c(2,2))
plot(model_w)
```

```{r}
results = data.frame(model_m_r,model_w_r,model_t_r)
data = c("error_rate","RMSE","R2")
rownames(results) <- data
results

df_coeff_a
df_coeff_w
df_coeff_m
```


