---
title: "Single and Multiple Linear Regression"
author: "Valeria Vera"
date: "2022-12-02"
output: html_document
---

```{r}
df <- read.csv('../data/all_data_merged_df(NEW).csv')
library(tidyverse)
library(ggpubr)
theme_set(theme_pubr())
library(ISLR2)
library(leaps)

library(tidyverse)
library(caret)
```

Linear regression of two variables
```{r}
head(df)
x = df$tot_men_enroll_2013
y = df$gradrate_ba_6yrswomen
model <- lm(x~y)
print(summary(model))
summary(model)$coefficient
#predict with the defined model
a <- data.frame(x = 2)
result <-  predict(model,a)
print(result)

plot(y,x,col = "blue",main = "",
     abline(model),cex = 1.3,pch = 16,xlab = "x",ylab = "y")
confint(model)

#residual standard error
sigma(model)/mean(x)

```
For a given the predictor, the t-statistic evaluates whether or not there is significant association between the predictor and the outcome variable, that is whether the beta coefficient of the predictor is significantly different from zero.

For a given predictor variable, the coefficient (b) can be interpreted as the average effect on y of a one unit increase in predictor, holding all other predictors fixed.



Multiple linear regression
```{r}
x = df$tot_men_enroll_2013
a = df$gradrate_ba_6yrswomen
b = df$gradrate_ba_6yrsblack_non.hispanic

model_mul <- lm(x ~ a + b )
print(summary(model_mul))
summary(model_mul)$coefficient


plot(x,a+b,col = "blue",main = "", abline(model_mul),cex = 1.3,pch = 16,xlab = "x",ylab = "y")

```
```{r}
anova(model, model_mul)
```
```{r}
ggplot(df, aes(x = off_campus_not_with_family_room_and_board, y =  gradrate_ba_6yrs_women)) +
  geom_point() +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 

model_mul <- lm(df$gradrate_ba_6yrs_women ~ df$off_campus_not_with_family_room_and_board)
print(summary(model_mul))
```
```{r}
ggplot(df, aes(x = off_campus_not_with_family_room_and_board, y =  gradrate_ba_6yrs_men)) +
  geom_point() +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 

ggplot(df, aes(x = off_campus_not_with_family_room_and_board, y =  gradrate_ba_6yrs_men)) +
  geom_point() +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
```
```{r}
ggplot(df, aes(x = off_campus_not_with_family_room_and_board, y =  gradrate_ba_6yrs_total)) +
  geom_point() +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 

model_mul <- lm(df$gradrate_ba_6yrs_women ~ df$off_campus_not_with_family_room_and_board)
print(summary(model_mul)$coefficient)
confint(model)
```

```{r}

#renaming columns
colnames(df)[which(names(df) == "gradrate_ba_6yrs_women")] <- "gradrate_women"
colnames(df)[which(names(df) == "gradrate_ba_6yrs_men")] <- "gradrate_men"
colnames(df)[which(names(df) == "percent_of_undergraduate_students_awarded_federal_state_local_institutional_or_other_sources_of_grant_aid")] <- "students_percent_aid"
colnames(df)[which(names(df) == "percent_of_undergraduate_students_awarded_pell_grants")] <- "students_percent_pellg"
colnames(df)[which(names(df) == "average_amount_of_federal_state_local_institutional_or_other_sources_of_grant_aid_awarded_to_undergraduate_students")] <- "students_avg_aid"
colnames(df)[which(names(df) == "average_amount_pell_grant_aid_awarded_to_undergraduate_students")] <- "students_avg_pellg"
colnames(df)[which(names(df) == "off_campus_not_with_family_room_and_board")] <- "living_exp_off_c"
colnames(df)[which(names(df) == "on_campus_room_and_board")] <- "living_exp_on_c"

df = df %>% drop_na(gradrate_women,gradrate_men,students_percent_aid,students_percent_pellg,students_avg_aid,students_avg_pellg,living_exp_off_c)
```

```{r}
y <- ggplot(df, aes(x = living_exp_on_c, y =  gradrate_women)) +
  geom_point(aes(color = living_exp_on_c),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and living expenses on campus") +
  xlab("living expenses by city (dollars)") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = living_exp_on_c, y =  gradrate_men)) +
  geom_point(aes(color = living_exp_on_c),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and living expenses on campus") +
  xlab("living expenses by city (dollars)") + ylab("Graduate rate of men")+ theme_gray() 

```


```{r}
y <- ggplot(df, aes(x = living_exp_off_c, y =  gradrate_women)) +
  geom_point(aes(color = living_exp_off_c),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and living expenses out campus") +
  xlab("living expenses by city (dollars)") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = living_exp_off_c, y =  gradrate_men)) +
  geom_point(aes(color = living_exp_off_c),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and living expenses out campus") +
  xlab("living expenses by city (dollars)") + ylab("Graduate rate of men")+ theme_gray() 

```
```{r}
y <- ggplot(df, aes(x = students_percent_aid, y =  gradrate_women)) +
  geom_point(aes(color = students_percent_aid),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_percent_aid") +
  xlab("students_percent_aid") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = students_percent_aid, y =  gradrate_men)) +
  geom_point(aes(color = students_percent_aid),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_percent_aid") +
  xlab("students_percent_aid") + ylab("Graduate rate of men")+ theme_gray() 

```

```{r}
y <- ggplot(df, aes(x = students_percent_pellg, y =  gradrate_women)) +
  geom_point(aes(color = students_percent_pellg),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_percent_pellg") +
  xlab("students_percent_pellg") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = students_percent_pellg, y =  gradrate_men)) +
  geom_point(aes(color = students_percent_pellg),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_percent_pellg") +
  xlab("students_percent_pellg") + ylab("Graduate rate of men")+ theme_gray() 

```

```{r}
y <- ggplot(df, aes(x = students_avg_aid, y =  gradrate_women)) +
  geom_point(aes(color = students_avg_aid),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_avg_aid") +
  xlab("students_percent_pellg") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = students_avg_aid, y =  gradrate_men)) +
  geom_point(aes(color = students_avg_aid),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_avg_aid") +
  xlab("students_avg_aid") + ylab("Graduate rate of men")+ theme_gray() 

```

```{r}
y <- ggplot(df, aes(x = students_avg_pellg, y =  gradrate_women)) +
  geom_point(aes(color = students_avg_pellg),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_avg_pellg") +
  xlab("students_percent_pellg") + ylab("Graduate rate of woman")+ theme_gray() 

y <- ggplot(df, aes(x = students_avg_pellg, y =  gradrate_men)) +
  geom_point(aes(color = students_avg_pellg),size=.8) +
  scale_color_gradient(low = "orange", high = "darkblue") +
  stat_smooth()+geom_smooth(method='lm', se=FALSE, color = "red", size=0.5,linetype = "dashed") 
y + ggtitle("Relation of graduate rate by gender and students_avg_pellg") +
  xlab("students_avg_pellg") + ylab("Graduate rate of men")+ theme_gray() 

```


```{r}
#separating data
set.seed(1)
sample <- sample(c(TRUE, FALSE), nrow(df_clean), replace=TRUE, prob=c(0.8,0.2))
train  <- df[sample, ]
test   <- df[!sample, ]


#define and train model
model_m<- lm(gradrate_men ~ (students_percent_aid+students_percent_pellg+
                             students_avg_aid+students_avg_pellg+
                             +living_exp_off_c)^2, data = train)
summary(model_m)

#evaluate model
error_rate = sigma(model_m)/mean(train$gradrate_men)
y_hat <- predict(model_m, test)
predictions <- model_m %>% predict(test)
RMSE(predictions, test$gradrate_men)
R2(predictions, test$gradrate_men)

par(mfrow=c(2,2))
plot(model_m)
```


```{r}

#separating data
set.seed(1)
sample <- sample(c(TRUE, FALSE), nrow(df_clean), replace=TRUE, prob=c(0.8,0.2))
train  <- df[sample, ]
test   <- df[!sample, ]

#define and train model
model_w<- lm(gradrate_women ~ (students_percent_aid+students_percent_pellg+
                             students_avg_aid+students_avg_pellg+
                             +living_exp_off_c)^2, data = train)
summary(model_w)

#evaluate model
error_rate = sigma(model_w)/mean(train$gradrate_women)
y_hat <- predict(model_w, test)
predictions <- model_w %>% predict(test)
RMSE(predictions, test$gradrate_women)
R2(predictions, test$gradrate_women)

par(mfrow=c(2,2))
plot(model_w)
```
```{r}
#anova(model_m,model_w)

#Model 1: value ~ weight + clarity
#Model 2: value ~ weight + clarity + color
 
```


