---
title: "sonali_hypothesis_testing"
author: "Sonali Dabhi"
date: "2022-11-30"
output: html_document
---

```{r}
df <- read.csv('../data/all_data_merged_df.csv')
library(tidyverse)
```

```{r}
# do some data cleaning to try to scope out outliers - these are the cases in which 6-year grad rates are 100% (might want to remove these)

#took the code from EDA code and removed these from the data set for hypothesis testing: 

id_cols <- c("unitid", "institution.name.x", "institution_entity_name_hd2020", "state.abbreviation_hd2020", "city.location.of.institution_hd2020",
             "zip.code_hd2020", "institutional.category_hd2020")
full_grad_rate <- df %>% subset(gradrate_ba_6yrstotal == 100) %>% dplyr::select(append(id_cols, colnames(df)[grep("grad(uation.)?rate", colnames(df))]))

df_subset  <- df %>% slice(-c(71,571, 621, 682, 686))

write.csv(df_subset, "../data/data_merged_df.csv", row.names = FALSE)
```



###Pell-Grant recipients are graduating at *higher* rates than the general student body?
Ho: There is no signification difference between the average graduation rates between pell grant recipients and the general student body. 

Ha: Pell grant recipients have a significantly lower average graduation rates than the general student body. 


```{r}
gen_body <- df_subset$gradrate_ba_6yrstotal
pell_rate <- df_subset$pell.grant.recipients_.overall.graduation.rate.within.150.percent.of.normal.time

t.test(pell_rate, gen_body, alt="less")

```

###Pell-Grant recipients are graduating at *higher* rates than the general student body based on  dimension of school size


Ho: There is no significant difference between Pell-Grant recipients graduation rates at small and larger schools

Ha:Pell-Grant recipients at smaller schools have higher graduation rates than those at larger schools
 
```{r}

```



###Women graduate at higher rates than men 

Ho: There is no signification difference between the average graduation rates between men and women

Ha: Women have a significantly higher average graduation rates than men.


```{r}
grad_wom <- df_subset$gradrate_ba_6yrswomen

grad_men <- df_subset$gradrate_ba_6yrsmen

t.test(grad_wom, grad_men, alt="greater")

```

###Black students graduate at higher rates than white students


Ho: There is no signification difference between the average graduation rates between black and white students.

Ha: Black students have a significantly lower average graduation rates than white students.


```{r}
bla_grad <- df_subset$gradrate_ba_6yrsblack_non.hispanic
nonbla_grad <- df_subset$gradrate_ba_6yrswhite_non.hispanic

t.test(bla_grad, nonbla_grad, alt="less")

```
###Hispanic students graduate at higher rates than white students

Ho: There is no signification difference between the average graduation rates between hispanic and non-hispanic students.

Ha: Hispanic students have a significantly lower average graduation rates than non-hispanic students.


```{r}
hisp_grad <- df_subset$gradrate_ba_6yrshispanic

all_nonhisp_grad <- append(df_subset$gradrate_ba_6yrswhite_non.hispanic,  nonbla_grad)

t.test(all_nonhisp_grad, nonhisp_grad, alt="less")




all_nonhisp_grad <-df_subset$gradrate_ba_6yrswhite_non.hispanic

t.test(hisp_grad, nonhisp_grad, alt="less")
```

