---
title: "sonali_hypothesis_testing"
author: "Sonali Dabhi"
date: "2022-11-30"
output: html_document
---

```{r}
df <- read.csv('../data/all_data_merged_df.csv')
df_1 <- read.csv('../data/all_data_merged_df(NEW).csv')
df_2 <- read.csv('../data/enrollement_completer.csv')
library(tidyverse)

#maybe remove colleges with no black/hisp/pell grant
```

```{r}
# do some data cleaning to try to scope out outliers - these are the cases in which 6-year grad rates are 100% (might want to remove these)
#took the code from EDA code and removed these from the data set for hypothesis testing: 

id_cols <- c("unitid", "institution_name", "undergraduate_enrollment", "percent_of_undergraduate_enrollment_that_are_black_or_african_american",              "percent_of_undergraduate_students_awarded_pell_grants")
full_grad_rate <- df_1 %>% subset(gradrate_ba_6yrs_total == 100) %>% dplyr::select(append(id_cols, colnames(df_1)[grep("grad(uation.)?rate", colnames(df_1))]))

df_subset  <- df_1 %>% slice(-c(485, 538))

#write.csv(df_subset, "../data/data_merged_df.csv", row.names = FALSE)
```



###Pell-Grant recipients are graduating at *higher* rates than the general student body?

Ho: There is no signification difference between the average graduation rates between pell grant recipients and the general student body. 

Ha: Pell grant recipients have a significantly lower average graduation rates than the general student body. 

```{r}
#gen_body is the general body population for each school
#pell_rate is looking at at those graduate rates for pell grant recipients 
gen_body <- df_subset$gradrate_ba_6yrs_total
pell_rate <- df_subset$pell_grant_recipients_overall_graduation_rate_within_150_percent_of_normal_time
t.test(pell_rate, gen_body, alt="less")
```

###Pell-Grant recipients are graduating at *higher* rates than the general student body based on dimension of school size

Ho: There is no significant difference between Pell-Grant recipients graduation rates at small and larger schools

Ha: Pell-Grant recipients at bigger schools have higher graduation rates than those at smaller schools
 
```{r}
enrollment <- df_subset$undergraduate_enrollment
#first we need to classify what is small vs large school, the easiest way would to be splitting based on the median
hist(enrollment)
#weird spread but ok 
split_val <- median(enrollment)
pell_small <- subset(df_subset$pell_grant_recipients_overall_graduation_rate_within_150_percent_of_normal_time, df_subset$undergraduate_enrollment < split_val )
pell_big <- subset(df_subset$pell_grant_recipients_overall_graduation_rate_within_150_percent_of_normal_time, df_subset$undergraduate_enrollment > split_val )

t.test(pell_big, pell_small, alt="greater")

```



###Women graduate at higher rates than men 

Ho: There is no signification difference between the average graduation rates between men and women

Ha: Women have a significantly higher average graduation rates than men.


```{r}
grad_wom <- df_subset$gradrate_ba_6yrs_women

grad_men <- df_subset$gradrate_ba_6yrs_men

t.test(grad_wom, grad_men, alt="greater")

```

###Black students graduate at higher rates than white students


Ho: There is no signification difference between the average graduation rates between black and white students.

Ha: Black students have a significantly lower average graduation rates than white students.


```{r}
bla_grad <- df_subset$gradrate_ba_6yrs_black_non_hispanic
nonbla_grad <- df_subset$gradrate_ba_6yrs_white_non_hispanic

t.test(bla_grad, nonbla_grad, alt="less")

```
###Hispanic students graduate at higher rates than white students

Ho: There is no signification difference between the average graduation rates between hispanic and non-hispanic students.

Ha: Hispanic students have a significantly lower average graduation rates than non-hispanic students.


```{r}
hisp_grad <- df_subset$gradrate_ba_6yrs_hispanic

all_nonhisp_grad <- append(df_subset$gradrate_ba_6yrs_white_non_hispanic,  bla_grad)

t.test(all_nonhisp_grad, hisp_grad, alt="less")
```
Chi-squared for independence: 

##Is there a relationship between race and graduation rates? 

As we are asking if they are independent, we should do a Chi-squared test for independence. 

Ho: There is no association between graduation rate and known graduate race in the population of interest. 

Ha: There is an association between graduation rate and known graduate race in the population of interest.



Chi-squared goodness of fit: 

##Does the graduate rate differ from enrollment rate

Ho: The graduation count data comes from the distribution of the enrollment data. 

Ha: The graduation count data does not come from the distribution of the enrollment data. 

```{r}
col_name_enrollment <- c("unitid","percent_of_undergraduate_enrollment_that_are_black_or_african_american", "percent_of_undergraduate_enrollment_that_are_hispanic_latino", "percent_of_undergraduate_enrollment_that_are_white","percent_of_undergraduate_enrollment_that_are_asian", "percent_of_undergraduate_enrollment_that_are_american_indian_or_alaska_native", "percent_of_undergraduate_enrollment_that_are_native_hawaiian_or_other_pacific_islander","undergraduate_enrollment","gradrate_ba_6yrs_black_non_hispanic","gradrate_ba_6yrs_hispanic","gradrate_ba_6yrs_white_non_hispanic", "gradrate_ba_6yrs_asian", "gradrate_ba_6yrs_american_indian_or_alaska_native", "gradrate_ba_6yrs_native_hawaiian_or_other_pacific_islander" ,"undergraduate_enrollment")
chi_df <- subset(df_subset, select = col_name_enrollment)


chi_df <- chi_df %>% mutate(count_white_enroll = undergraduate_enrollment * percent_of_undergraduate_enrollment_that_are_white/100)
chi_df <- chi_df %>% mutate(count_black_enroll = undergraduate_enrollment * percent_of_undergraduate_enrollment_that_are_black_or_african_american/100)
chi_df <- chi_df %>% mutate(count_hisp_enroll = undergraduate_enrollment * percent_of_undergraduate_enrollment_that_are_hispanic_latino/100)
chi_df <- chi_df %>% mutate(count_asian_enroll = undergraduate_enrollment * percent_of_undergraduate_enrollment_that_are_asian/100)
chi_df <- chi_df %>% mutate(count_ask_enroll = undergraduate_enrollment * percent_of_undergraduate_enrollment_that_are_american_indian_or_alaska_native/100)
chi_df <- chi_df %>% mutate(count_haw_enroll = undergraduate_enrollment * percent_of_undergraduate_enrollment_that_are_native_hawaiian_or_other_pacific_islander/100)


chi_df <- chi_df %>% mutate(count_white_grad = count_white_enroll * gradrate_ba_6yrs_white_non_hispanic/100)
chi_df <- chi_df %>% mutate(count_black_grad = count_black_enroll * gradrate_ba_6yrs_black_non_hispanic/100)
chi_df <- chi_df %>% mutate(count_hisp_grad = count_hisp_enroll * gradrate_ba_6yrs_hispanic/100)
chi_df <- chi_df %>% mutate(count_asian_grad = count_asian_enroll * gradrate_ba_6yrs_asian/100)
chi_df <- chi_df %>% mutate(count_ask_grad = count_ask_enroll * gradrate_ba_6yrs_american_indian_or_alaska_native/100)
chi_df <- chi_df %>% mutate(count_haw_grad = count_haw_enroll * gradrate_ba_6yrs_native_hawaiian_or_other_pacific_islander/100)

exp_count <- c(mean(chi_df$count_white_enroll), mean(chi_df$count_black_enroll), mean(chi_df$count_hisp_enroll), mean(chi_df$count_asian_enroll), mean(chi_df$count_ask_enroll), mean(chi_df$count_haw_enroll))

obs_count <- c(mean(chi_df$count_white_grad), mean(chi_df$count_black_grad), mean(chi_df$count_hisp_grad),mean(chi_df$count_asian_grad), mean(chi_df$count_ask_grad),mean(chi_df$count_haw_grad) )

chidf <- t(data.frame(exp_count, obs_count))

chisq <- chisq.test(chidf)
chisq

exp_chi <- chisq.test(chidf)$expected
print(exp_chi)


#yates correction:
chisq_1 <- chisq.test(chidf, correct = TRUE, simulate.p.value = TRUE)
chisq_1
```

As our p value is less that 0.05, we can reject the null hypothesis. There are different counts between the graduation and enrollment counts. 



##ANOVA:
Ho: There is no difference among group means. 

Ha: There is a difference between the group means. (One group differs significantly from the overall mean of the dependent variable) 

```{r}
col_name_anova <- c("unitid", "sector_of_institution", "gradrate_ba_6yrs_total", "diversity_quantiles", "undergraduate_enrollment")
anova_df <- subset(df_subset, select = col_name_anova)
anova_df <- anova_df %>% mutate(count_grad = log(undergraduate_enrollment * gradrate_ba_6yrs_total/100))


hist(log(anova_df$count_grad))
#would be doing a two way ANOVA: 

two.way <- aov(count_grad ~ sector_of_institution + diversity_quantiles, data = anova_df)
summary(two.way)

interaction <- aov(count_grad ~ sector_of_institution + diversity_quantiles + sector_of_institution*diversity_quantiles, data = anova_df)
summary(interaction)

```

Understanding results: 
- These rows inform us whether our independent variables and their interaction have a statistically significant effect on the dependent variable.

There are statistically significant differences between sector, diversity, and their interaction on the log of graduation counts.



